---
title: "Final"
author: "Abdullah Akkoloğlu"
date: "2024-01-12"
output: html_document
---

```{r}
knitr::opts_chunk$set(warning = FALSE)
```

```{r}
# Set the CRAN mirror
options(repos = c(CRAN = "https://cran.r-project.org"))
#install.packages("caret")
#install.packages("ggcorrplot")
#install.packages("corrr")
#install.packages("ggplot2")
#install.packages("corrplot")
#install.packages("factoextra")
#install.packages("FactoMineR")
library("FactoMineR")
library(corrplot)
library(ggplot2)
library('corrr')
library(ggcorrplot)
library(caret)
library("factoextra")
```
# 1. (5p) Please find your original dataset or datasets; and describe your data in the first step.

```{r}
rawData <- read.csv("nba.csv")
set.seed(123)
```
# 2. (4p) Use “Exploratory data analysis”. Write down your comments.

```{r}
head(rawData)
summary(rawData)
```
# 3. (4p) Use some “visualization techniques” and talk about your data further.
```{r}
ggplot(rawData, aes(x=PTS)) + geom_histogram()
ggplot(rawData, aes(x=MIN, y=PTS)) + geom_point()
ggplot(rawData, aes(x=factor(TARGET_5Yrs))) + geom_bar()
```
# KNN FOR NA VALUES
```{r}
data <- rawData[, -which(names(rawData) == "Name")]
sum(is.na(data))
preProcess_missingdata_model <- preProcess(data, method='knnImpute')
preProcess_missingdata_model
data <- predict(preProcess_missingdata_model, newdata = data)
sum(is.na(data))
data$TARGET_5Yrs <- ifelse(data$TARGET_5Yrs < 0, 0,
                                   ifelse(data$TARGET_5Yrs > 0, 1,
                                          data$TARGET_5Yrs))
```
# (4p) Check your data for multicollinearity, make your comments.
```{r}
independent_vars <- data[, names(data) != "TARGET_5Yrs"]
data_scaled <- scale(independent_vars)
data_scaled <- data.frame(data_scaled, target = data$TARGET_5Yrs)

cor_matrix <- cor(data_scaled)
corrplot(cor_matrix)
```
# PCA:
```{r}
pca_result <- princomp(data_scaled, cor = TRUE)
summary(pca_result)
fviz_eig(pca_result, addlabels = TRUE)
fviz_pca_var(pca_result, col.var = "black")
pcadata <- pca_result$scores[,1:5]
pcadata  <- as.data.frame(pcadata)
pcadata$target <- data_scaled$target
```
# LOG REG ORIGINAL:
```{r}
ind<-sample(2,nrow(data_scaled),replace=T,prob = c(0.7,0.3))
train<-data_scaled[ind==1,]
test<-data_scaled[ind==2,]
fit<-glm(target~.,family=binomial,train)
train$probs<-predict(fit, type = 'response')
train$predict<-rep('0',945)
train$predict[train$probs>0.5]<-"1"
table(train$predict,train$target)
confusionMatrix(as.factor(train$predict), as.factor(train$target))

fit<-glm(target~.,family=binomial,test)
test$probs<-predict(fit, type = 'response')
test$predict<-rep('0',395)
test$predict[test$probs>0.5]<-"1"
table(test$predict,test$target)
confusionMatrix(as.factor(test$predict), as.factor(test$target))
```
# LOG REG WITH PCA:
```{r}
ind<-sample(2,nrow(pcadata),replace=T,prob = c(0.7,0.3))
train_pca<-pcadata[ind==1,]
test_pca<-pcadata[ind==2,]
fit_pca<-glm(target~.,family=binomial,train_pca)
train_pca$probs<-predict(fit_pca, type = 'response')
train_pca$predict<-rep('0',945)
train_pca$predict[train_pca$probs>0.5]<-"1"
table(train_pca$predict,train_pca$target)
confusionMatrix(as.factor(train_pca$predict), as.factor(train_pca$target))

test_pca$prob<-predict(fit_pca,newdata = test_pca, type = 'response')
test_pca$predict<-rep('0',395)
test_pca$predict[test_pca$prob>0.5]<-"1"
table(test_pca$predict,test_pca$target)
confusionMatrix(as.factor(test_pca$predict), as.factor(test_pca$target))
```
# HIERARCHICAL CLUSTERING
```{r}
data_scaled <- scale(pcadata[, -which(names(pcadata) == "target")])
dist_matrix <- dist(data_scaled, method = "euclidean")
hc <- hclust(dist_matrix, method = "ward.D2")
plot(hc)
cut_height <- 10
clusters <- cutree(hc, h = cut_height)
table(clusters)
data_with_clusters <- cbind(pcadata, Cluster = clusters)

for(i in unique(data_with_clusters$Cluster)){
    cat("Cluster", i)
    cluster_table <- table(data_with_clusters[data_with_clusters$Cluster == i, "target"])
    cluster_percent <- prop.table(cluster_table) * 100  
    print(cluster_table)
    print(cluster_percent)
    cat("\n")
}


```
# K-MEANS CLUSTERING
```{r}
data_scaled <- scale(pcadata[, -which(names(pcadata) == "target")])
wss <- (nrow(data_scaled)-1)*sum(apply(data_scaled,2,var))
for (i in 2:15) {
     wss[i] <- sum(kmeans(data_scaled, centers=i)$withinss)
}
plot(1:15, wss, type="b")
set.seed(123)
kmeans_result <- kmeans(data_scaled, centers=8)
table(kmeans_result$cluster, pcadata$target)
```

```{r}

```