---
title: "Mole"
author: "Abdullah Akkoloğlu"
date: "2023-12-18"
output: html_document
---

```{r}
# Set the CRAN mirror
options(repos = c(CRAN = "https://cran.r-project.org"))

# Install the packages
install.packages("caret")
install.packages("skimr")

# Load the libraries
library(caret)
library(skimr)
library(RANN)
```
Read the data.
```{r}
mole <- read.csv("mole2_dataset.csv")
```

1.Split your data into train (80%) and test (20%) datasets using caret’s createDataPartition function.
```{r}
set.seed(100)
trainRowNumbers <- createDataPartition(mole$diagnosis, p=0.8, list=FALSE)
trainData <- mole[trainRowNumbers,]
testData <- mole[-trainRowNumbers,]
```

2 Use the skim_to_wide function in skimr package, and provide descriptive stats for each column. Comment!
```{r}
skimmed <- skim_to_wide(trainData)
skimmed
```

3 Predict and impute the missing values with k-Nearest Neighbors using preProcess function in caret. Comment!
```{r}
preProcess_missingdata_model <- preProcess(trainData, method='knnImpute')
preProcess_missingdata_model
trainData <- predict(preProcess_missingdata_model, newdata = trainData)
```

4. After you impute missing values, use variable transformations. Convert all the numeric variables to range between 0 and 1, by setting method=range in preProcess function.
```{r}
preProcess_range_model <- preProcess(trainData, method='range')
trainData <- predict(preProcess_range_model, newdata = trainData)
```

5. Use caret’s featurePlot() function to visually examine how the predictors influence the predictor variable. Comment!
basel cell carcinoma - melanoma - nevus - seborrheic keratosis
```{r}
trainData$diagnosis <- as.factor(trainData$diagnosis)
featurePlot(x = trainData[, 6:13], 
            y = trainData$diagnosis, 
            plot = "box",
            strip=strip.custom(par.strip.text=list(cex=.7)),
            scales = list(x = list(relation="free"), 
                          y = list(relation="free")))
```










